PoVaTiM: A Possibilistic Valid-Time Model \\ \vspace*{0.04truein}

### 2012-06-22 ijcis-pos-model/sources/temporaldatabases.tex
last updates

%	\begin{itemize}	
%	\item \textbf{Granularity} denotes a partition on the set of chronons. The conversion between several granularities is studied in \cite{Lin97efficientconversion}. Granularity is the base of the temporal model in \cite{Cru97}. An object-oriented implementation is in \cite{624013}.
%	\item	To ensure \textbf{consistency}, temporal databases usually redefine the primary key of a relation. The new primary key takes into account the presence of the time. In order to keep the consistency, the DML is re-defined. For example, in a VTDB, the temporal update sentence is usually composed by an update sentence (closing the old row) plus an insert sentence (creating the new row).
%	\end{itemize}
%Guy's suggestion: instead of imprecision, use imperfection which is more general.
\subsubsection{Imperfection and time}
Representing imprecision and its semantics when dealing with time has been studied for a long time. Several proposals for representing and computing imprecise time indications can be found in \cite{DeCaluwe1997,DeTre1997}. Also, the changes between several granularities can be seen as a source of imprecision \cite{Devos1998}.

In the proposal section we will consider two kinds of imprecision:
\begin{itemize}
\item \textbf{Imperfection in the database} denotes the uncertainty that arises when the knowledge about the temporal data in the database is uncertain. E.g., a database record shows that \emph{`The car is in the garage around April.'}
 \item \textbf{Imprecision in the query specification} denotes the imprecision in the specification of temporal criteria by the user, when querying. E.g., \emph{`The user wants to obtain a car which is red and which is in the garage around April.'}
\end{itemize}

\subsubsection{Representation}
Several proposals for managing uncertain time in a database exist. Some proposals work with rough sets \cite{Qiang2009}, other proposals rely on possibility distributions for representing uncertainty in time \cite{Dyreson1998,Garrido2009,Galindo2001}. In order to compare temporal possibility distributions, extensions of the classical Allen's operators \cite{Allen1983} are defined in \cite{Ohlbach2004,Nagypal2003,Dubois2003a,Schockaert2008}.
%In the proposal section, we will follow the representation by means of possibility distributions, in order to work with both satisfaction and dissatisfaction degrees. Also, in order to work properly with fuzzy operators, the underlying domain should be numeric. 
% Take a look into this:
%In this paper, the representation for the dates will follow the Julian Day Number (JDN) representation \cite{Dir96}.

@@

In order to deal with uncertainty in time intervals, several proposals are made. Here, two approaches are introduced: the first one, based on \emph{Fuzzy Validity Period}~\cite{Garrido2009} and the second one based on \emph{Possibilistic Valid-time Period}~\cite{JoseEnriquePons2012}.

\begin{definition}
A \textbf{Fuzzy Validity Period} (\emph{FVP}) is defined as a fuzzy time interval specifying when the data regarding an object is valid. A fuzzy time interval is then the fuzzification of a crisp time interval.
\end{definition}
Several options to transform possibility distributions corresponding to the fuzzy starting point and the fuzzy end point into one consistent FVP exist \cite{Garrido2009}, e.g (Fig. \ref{fig:fuzzy-validity-period}):
\begin{itemize}
\item The \textbf{convex hull} approach is the most intuitive approach. The resulting FVP is the convex hull of the union of both fuzzy sets.
\item The \textbf{uncertainty preserving} approach is less intuitive but more realistic. The amount of uncertainty is maintained at the edges of the possibility distribution representing the FVP \cite{Garrido2009}.
\end{itemize}

%%%%%%
% FUZZY VALIDITY PERIOD
%%%%%%
\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=0.25]{./graphs/comparisoncv.pdf}

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:fuzzy-validity-period}Transformation to obtain the FVP. The top graph shows the two triangular possibility distributions. The middle graph shows the convex hull validity period, the bottom one shows the result of the second transformation, which maintains the imprecision.}
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}

\begin{definition}
A \textbf{Possibilistic Valid-time Period} (\emph{PVP}) is an ill-known interval in time specifying when the data regarding an object is valid.
\end{definition}
A PVP is an ill-known interval in the sense of definition \ref{def;possibilistic-variable}, section \ref{subsec:possibilistic-variables}. Note that this representation is \emph{disjunctive}: the PVP represents only one crisp time interval, but that for some reason it is (partially) unknown.

The ill-known interval approach has many advantages from the representation by the FVP as demonstrated in \cite{Pons2011},\cite{Pons2012}. Table \ref{tbl:comparative-pvp-fvp} is a comparative between PVP and FVP. The following list defines the items in the comparative:

\begin{enumerate}
\item Domain: The domain of the possibility distribution modelled by the approach.
\item Implementation of relationships: How to implement a relationship.
\item Allen's relations: Are the Allen's relations defined?
\item Storage: The way the data is stored in the database.
\item Possibility measures: Does the framework provides always a possibility measure for any relation between the temporal elements?
\item Necessity measures:  Does the framework provides always a necessity measure for any relation between the temporal elements?
\end{enumerate}

%%
%% comparison table between pvp and fvp.
%%
\vglue13pt
%\begin{table}[htbp]
\tcap{Comparative PVP vs FVP}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c p{2cm} p{2cm}}\\
Item & PVP & FVP\\
\hline
(1) & $\Pow(\R)$ & $\R$ \\
(2) & Ill-known constraints. & Ad-hoc operators. \\
(3) & $\checkmark$ & - \\
(4) & Two distributions one for each endpoint. & Only one distribution. \\ 
(5) & $\checkmark$ & $\checkmark$ \\
(6) & $\checkmark$ & - \\
\hline\\
\end{tabular}
\label{tbl:comparative-pvp-fvp}
}

In the rest of the paper we will work only with PVP to represent valid-time intervals.

\subsection{\label{subsubsec:Understanding-valid-time-databases}Understanding Valid-time Databases}
This subsection is devoted to describe the behaviour of a crisp valid-time database. For the shake of simplicity, only the three main operations in the Data Manipulation Language \emph{DML} a.k.a \emph{CRUD} (CReate Update, Delete) operations are shown. Usually the DML operations in a temporal database are re-defined e.g. a typical update sentence in SQL could be re-defined by means of several insert and update sentences. Therefore, in order to be clear, these high level primitives in the DML for a valid-time database are usually noted as Insert, Modify and Delete. In the following subsections each primitive is defined and explained. Finally a illustrative example is given. For a more complete information on the behaviour of a bitemporal database, please refer to ~\cite{Jensen1994}.

\begin{definition}
Consider a relation $R$ in a temporal database, an object $A$ given by the attributes $\left(a_1, \ldots, a_n \right)$ and the crisp time interval $I$ given by $\left[a,\ b\right] : a,b \in \T, a \leq b$. A valid-time object $A$ in the relation $R$ is specified by:
\begin{equation}
\label{eq:rel-def}
\left( R, A , I \right)
\end{equation}
\end{definition}

In order to simplify the algorithms for the manipulation of data, some auxiliary functions are defined:

\begin{definition}
\label{def:close-a-crisp-interval-r}
Consider two crisp intervals $I= \left[a ,b \right]$ and $J= \left[c ,d \right]$, $c > a$. The CloseR$\left(I, J\right)$ function allows to close the right-open interval $I$ with respect to the first value $d$ in $J$:
\begin{align}
\label{eq:close-a-crisp-interval}
\mbox{CloseR} \left( I, J \right) &=& \\ 
\begin{cases}
\nonumber
I & \mbox{ if } b \neq +\infty \\
I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
\end{cases}
\end{align}
\end{definition}

@@

Now it is possible to close the current version of an object by using \eqref{def:close-a-crisp-interval-r} and \eqref{eq:current-in-relation}

\begin{definition}
\label{def:close-current-version}
Consider an object $A$, the relation $R$ and a time interval $J$. Then, the function close-current$\left(R,A,J \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J \right)$.

\begin{eqnarray}
\label{eq:close-current}
\text{Close-current} \left(R, A, J \right) =\\
\begin{cases}
\nonumber
\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
\nonumber
\mbox{ if } I = \mbox{ Current } \left(R, A \right) \wedge J > \mbox{ CloseR } \left(I, J \right)   \\
\nonumber R , \text{ in any other case}
\end{cases}
\end{eqnarray}
\end{definition}

%It is also possible to append at the beggining by using \eqref{def:close-a-crisp-interval-l} and \eqref{def:first-in-relation}.
%\begin{definition}
%\label{def:append-first-version}
%Consider an object $A$, the relation $R$ and a time interval $J$. Then, the function append-first$\left(R,A,J \right)$ closes any right open version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
%
%\begin{eqnarray}
%%\label{eq:close-current}
%\text{Append-first} \left(R, A, J \right) =\\
%\begin{cases}
%\nonumber
%\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseL } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
%\nonumber
%\mbox{ if } I = \mbox{ First } \left(R, A \right) \wedge J < \mbox{ CloseL } \left(I, J \right)   \\
%\nonumber R , \text{ in any other case}
%\end{cases}
%\end{eqnarray}
%\end{definition}

\subsubsection{\label{subsubsec:insert}Insert}
The user wants to store an object $A$ which is valid in the relation $R$ during the time interval specified by $I = \left(a, b \right)$.
%
%
%The interpretation of the insert operation is the following: The user wants to store in the database that an object $A$  is true for some period(s) of time $I$ in the given relation $R$. To indicate that the object $A$ is current in the relation, the special value Until Current, $\UC$  is used. 
There are two main cases when performing a create operation:
\begin{enumerate}
\item The object $A$ was never in the relation $R$: The object is added with the valid-time indicated by the crisp interval $I$.

\item The object $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are two possibilities:
	\begin{enumerate}
	\item Insert $\left(A, I\right)$ in the relation $R$. If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
	\item Reject the insertion, if the time interval $I$ do overlap any existing valid-time interval for the object $A$ in the relation $R$.
	\end{enumerate}

>>>>

%	\begin{itemize}	
%	\item \textbf{Granularity} denotes a partition on the set of chronons. The conversion between several granularities is studied in \cite{Lin97efficientconversion}. Granularity is the base of the temporal model in \cite{Cru97}. An object-oriented implementation is in \cite{624013}.
%	\item	To ensure \textbf{consistency}, temporal databases usually redefine the primary key of a relation. The new primary key takes into account the presence of the time. In order to keep the consistency, the DML is re-defined. For example, in a VTDB, the temporal update sentence is usually composed by an update sentence (closing the old row) plus an insert sentence (creating the new row).
%	\end{itemize}
%Guy's suggestion: instead of imprecision, use imperfection which is more general.
\subsubsection{Imperfection and time}
The representation of imprecision and its semantics when dealing with time has been studied for a long time. Several proposals for representing and handling imprecise time indications can be found in \cite{DeCaluwe1997,DeTre1997}. Also, the changes between several granularities can be seen as a source of imprecision \cite{Devos1998}.

In this paper we will consider two kinds of imperfection:
\begin{itemize}
\item \textbf{Imperfection in the database}; the knowledge about the temporal data contains some imperfection. E.g., a database record shows that \emph{`The car was in the garage around April.'}
 \item \textbf{Imprecision in the query specification}; it denotes the imprecision in the specification of temporal criteria by the user, when querying. E.g., \emph{`The user wants a car which was in the garage around April.'}
\end{itemize}

\subsubsection{Representation}
Several proposals for managing uncertain time in a database exist. Some proposals work with rough sets \cite{Qiang2009}, and some others rely on possibility distributions for representing uncertainty in time \cite{Dyreson1998,Garrido2009,Galindo2001}. In order to compare temporal possibility distributions, the extensions of the classical Allen's operators \cite{Allen1983} are defined in \cite{Ohlbach2004,Nagypal2003,Dubois2003a,Schockaert2008}.
%In the proposal section, we will follow the representation by means of possibility distributions, in order to work with both satisfaction and dissatisfaction degrees. Also, in order to work properly with fuzzy operators, the underlying domain should be numeric. 
% Take a look into this:
%In this paper, the representation for the dates will follow the Julian Day Number (JDN) representation \cite{Dir96}.

@@

In order to deal with uncertainty in time intervals, several proposals have been made. Here, two approaches are described: the first one, based on \emph{Fuzzy Validity Periods}~\cite{Garrido2009} and the second one, based on \emph{Possibilistic Valid-time Periods}~\cite{JoseEnriquePons2012}.

\begin{definition}
A \textbf{Fuzzy Validity Period}\cite{Garrido2009} (\emph{FVP}) is defined as a fuzzy time interval specifying when the data regarding an object are valid. A fuzzy time interval is then the fuzzification of a crisp time interval.\\
\end{definition}

Several options to transform possibility distributions corresponding to the fuzzy starting point and the fuzzy end point into a consistent FVP exist \cite{Garrido2009}, e.g (Fig. \ref{fig:fuzzy-validity-period}):
\begin{itemize}
\item The \textbf{convex hull} approach is the most intuitive approach. The resulting FVP is the convex hull of the union of both possibility distributions.
\item The \textbf{uncertainty preserving} approach. The amount of uncertainty is maintained at the edges of the possibility distribution representing the FVP \cite{Garrido2009}.
\end{itemize}

%%%%%%
% FUZZY VALIDITY PERIOD
%%%%%%
\begin{samepage}
\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=0.25]{./graphs/comparisoncv.pdf}

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\end{samepage}

\fcaption{\label{fig:fuzzy-validity-period}Transformation to obtain the FVP. The top graph shows the two triangular possibility distributions. The middle graph shows the convex hull validity period, the bottom one shows the result of the second transformation, which maintains the imprecision.}
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}

The main feature for the FVP is the optimization for the storage. The compact representation is the result of a conjunctive semantic. The object is valid within all the time points inside the starting and ending points.

\begin{definition}
A \textbf{Possibilistic Valid-time Period} (\emph{PVP}) is an ill-known interval of time specifying when the data regarding an object are valid.
\end{definition}
A PVP is an ill-known interval in the sense of definition \ref{def;possibilistic-variable}. Note that the PVP represents only one crisp time interval, but for some reason, it is (partially) unknown.

The main feature for the PVP is that it preserves all the information for both starting and ending points\cite{Pons2011,Pons2012}. Table \ref{tbl:comparative-pvp-fvp} is a comparative between PVP and FVP. The following list defines the items in the comparative:

\begin{enumerate}
\item Domain: The domain of the possibility distribution modelled by the approach.
\item Implementation of relationships: How to implement a relationship.
\item Allen's relations: Are the Allen's relations defined?
\item Storage: The way the data are stored in the database.
\item Possibility measures: Does the framework provide always a possibility measure for any relation between the temporal elements?
\item Necessity measures:  Does the framework provide always a necessity measure for any relation between the temporal elements?
\end{enumerate}

%%
%% comparison table between pvp and fvp.
%%
\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{Comparative PVP vs FVP}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c p{2cm} p{2cm}}\\
Item & PVP & FVP\\
\hline
(1) & $\Pow(\R)$ & $\R$ \\
(2) & Ill-known constraints. & Ad-hoc operators. \\
(3) & $\checkmark$ & - \\
(4) & Two distributions one for each endpoint. & Only one distribution. \\ 
(5) & $\checkmark$ & $\checkmark$ \\
(6) & $\checkmark$ & - \\
\hline\\
\end{tabular}
\label{tbl:comparative-pvp-fvp}
} 
\end{samepage}
In the rest of the paper we will work only with PVP to represent valid-time intervals.

\subsection{\label{subsubsec:Understanding-valid-time-databases}Understanding Valid-time Databases}
This subsection is devoted to describe the behaviour of a crisp valid-time database. For the sake of simplicity, only the three main operations in the Data Manipulation Language (CReate Update, Delete) are shown. Usually the DML operations in a temporal database are re-defined (a typical update sentence in SQL could be expressed by means of a couple of insert and update sentences). Therefore, for the sake of clarity, these high level primitives in the DML for a valid-time database are usually noted as Insert, Modify and Delete. In the following subsections each primitive is defined and explained. Finally, an illustrative example is given. For a more complete information on the behaviour of a bitemporal database, please refer to ~\cite{Jensen1994}.

\begin{definition}
\label{def:valid-time-relation}
\emph{Valid-time relation.}
Consider the following definitions and notations:

\begin{itemize}
 \item A set of non-temporal attributes.
	\begin{equation}
	\label{eq:attribute-set}
	A = \left \lbrace A_1, A_2, \ldots, A_n \right \rbrace
	\end{equation}
      The domain for each attribute $A_1, \ldots, A_n$ is $D_1, \ldots, D_n$ respectively. 
\item The primary key for the attributes in $A$ is given by $A_K$:
      \begin{equation}
       \label{eq:primary-key-a}
      A_K \subset A
      \end{equation}
\item A set of two attributes; $S$  and $E$ the attributes for the starting and ending points respectively. $I$ defines the valid time interval for the data. 
\begin{equation}
 \label{eq:attribute-time-interval}
I = \left \lbrace S, E \right \rbrace
\end{equation}
\item Then $R$, the schema for the valid-time relation is:
\begin{equation}
 \label{eq:valid-time-relation}
R = A \cup  I
\end{equation}
\item The primary key for the valid-time relation $R$ is:
\begin{equation}
 \label{eq:valid-time-temporal-pk}
PK = A_K \cup I
\end{equation}

\item We will note by $r$ any valid instance of $R$. 
      \begin{equation}
       \label{eq:valid-time-instance}
      r \subseteq D_1\ x\ \ldots\ x\ D_n
      \end{equation}

\item Let $t$ be a tuple in the instance $r$; $t \in r$:
      \begin{itemize}
      \item We will note the values for the starting and the ending  points,$s$ and $e$  respectively. The value for the time interval is given by $i$.
      \begin{align}
       \label{eq:starting-point}
      s = t\left[S \right]\\
      e = t\left[E \right]\\
      i = \left(s, e\right)
      \end{align}

      \item Let $A_k \subset A$ be the set of the non-temporal attributes that are part of the primary key (equation \eqref{eq:primary-key-a}). Then, $a_k$ denotes the values for the non-temporal attributes of the primary key.
	    \begin{equation}
	     \label{eq:all-the-versions} 
	      a_k = t\left[A_K \right]
	    \end{equation}

    \item Let $PK$ be the primary key for the valid-time relation as given in equation \eqref{eq:valid-time-temporal-pk}. Then, $pk$ denotes the values for the attributes in the primary key.
	  \begin{equation}
	   \label{eq:value-pk}
	  pk = t\left[PK \right]
	  \end{equation}

      \end{itemize}
\item $Vt$ is the set with all the versions for a given tuple $t$.

\begin{equation}
 \label{eq:all-the-versions}
Vt = r\left(t\left[A_K\right] \right)
\end{equation}
If a tuple $t$ contains more than one version, the result of $Vt$ is a set. We will use the index $k$ to address the elements of the set. E.g. $Vt = \left \lbrace t_1, t_2, \ldots, t_n \right \rbrace$. Then $t_k , k \in \left \lbrace 1, \ldots, n \right \rbrace$ is an element of $Vt$.
\end{itemize}
\end{definition}
We will illustrate the definitions with an example.

\begin{example}
Consider the set of attributes $A = \left \lbrace A_1, A_2, A_3 \right \rbrace$. The primary key for these attributes is given by $A_K = \left \lbrace A_1, A_2 \right \rbrace$. Let $I = \left \lbrace S, E \right \rbrace$ be the set of temporal attributes that define the validity period of the data. $R$ is the valid-time relation and $r$ is an instance of the relation. The instance $r$ is given by the following elements. $r = \left \lbrace \left(a_{11}, a_{12}, a_{13}, s_1 ,e_1 \right) \right.$,  $\left(a_{21}, a_{22}, a_{23}, s_2, e_2 \right)$ , $\left. \left(a_{11}, a_{12}, a_{31}, s_3, e_3 \right) \right \rbrace$. The instance $r$ is illustrated in Table \ref{tbl:sample-definitions}. 
Consider the tuple $t = \left(a_{11}, a_{12}, a_{13}, s_1 ,e_1 \right)$. Then,
\begin{align}
 \nonumber
s_1 &= t\left[S \right]\\
 \nonumber
e_1 &= t[E]\\
 \nonumber
i\ \ &= \left(s_1, e_1\right)\\
 \nonumber
a_k &= t\left[A_K\right] = \left(a_{11}, a_{12} \right) \\
 \nonumber
pk &= t\left[PK\right] =\left(a_{11}, a_{12}, s_1, e_1\right)\\
 \nonumber
Vt &= r\left(t\left[A_K\right] \right) = \left \lbrace t_1, t_3 \right \rbrace
\end{align}

\end{example}

\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:sample-definitions}Sample database containing the instance $r$ of the valid-time relation $R$.}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c c c}\\
\hline
& \textbf{A$_1$}  & \textbf{A$_2$}  & $A_3$ & $S$ & $E$ \\
\hline
$t_1$&$a_{11}$ & $a_{12}$ & $a_{13}$ & $s_1$ & $e_1$ \\
$t_1$ & $a_{21}$ & $a_{22}$ & $a_{23}$ &  $s_2$ & $e_2$ \\
$t_1$ & $a_{11}$ & $a_{12}$ & $a_{31}$ & $s_3$ & $e_3$\\
\hline\\
\end{tabular}
}
\end{samepage}

% Consider $\A$ be the set of all the entities, and let $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity. Let $I = \left(S_i, E_i \right)$ be the set of attributes defining the time interval. The set $\I$ is the set of all the time intervals. The pair $\left(A, I\right)$ expresses that the data regarding the entity A are valid during the crisp time interval I. Let $R \subseteq \A \  x\  \I$ be a valid-time relation. Then, the following equation indicates that the pair $\left(A, I\right)$ is in the relation R:
% 
% \begin{equation}
% \label{eq:rel-def}
% \left( R, \left(A , I\right) \right)
% \end{equation}
% 
% 
In order to simplify the algorithms for the manipulation of data, some auxiliary functions and constants are defined:

\begin{definition}
 \label{def:from-beginning}
\emph{From the beginning (FB).} 
Consider the elements in definition \ref{def:valid-time-relation}. We will say $i = \left(FB, e \right)$ when $i = \left(- \infty,e \right)$. 
% $i = \left(s, e \right)$ be the value for a time interval as given equation \eqref{}.
% $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $I = \left(S_i, E_i \right), S_i, E_i \in \T$ the set of attributes defining the time interval. Let $R \subseteq \A\  x\  \I$ be a valid-time relation. We will say $I = \left(FB, E_i \right)$ when $I = \left(- \infty, E_i \right)$.
%The from the beginning constant is used when the value for $s_i$ is equal to $-\infty$.
\end{definition}

\begin{definition}
 \label{def:until-changed}
\emph{Until changed (UC).} 
Consider the elements in definition \ref{def:valid-time-relation}. We will say  $i = \left(s, UC \right)$ when $i = \left(s, +\infty \right)$.
% Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity. Let $\T$ be the time domain, and $I = \left(S_i, E_i \right), S_i, E_i \in \T$ the set of attributes defining the time interval. Let $R \subseteq \A\  x\  \I$ be a valid-time relation. We will say  $I = \left(S_i, UC \right)$ when $I = \left(S_i, +\infty \right)$.
%The until changed constant is used when the value for $e_i$ is equal to $+\infty$.
\end{definition}

\begin{definition}
\emph{Current.} 
Consider the elements in definition \ref{def:valid-time-relation}. We will say that the tuple $t$ is current in the instance $r$ of the relation $R$ when $i = \left(s, UC \right)$.
\end{definition}

%If the pair $\left(A, I \right)$ with $I = \left(S_i, UC \right), S_i \in \T$ is in the relation $R$, then, it is said that the pair $\left(A, I \right)$ is \emph{current} in the relation.

For example, let us consider the last row in Table \ref{tbl:library-sample}. The value for the time interval is $I = \left(4/4/2012 , UC \right)$. The meaning is that the patient with ID = 3 was in the hospital the  4/4/2012 and he or she is still there. The patient with ID = 3 is current in the relation.

\begin{example}
\label{ex:library-database}
 Consider a hospital database. Each patient is stored in the database with an identifier (ID), the name and the dates when the patient entered and left the hospital. Table \ref{tbl:library-sample} illustrates the first version of the database, after three insertions. In this example, $A = ($ ID, Name $)$ and $I = ($ Arrival, Leave $)$.
\end{example}

\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:library-sample}Sample hospital database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c}\\
\hline
\textbf{ID}  & Name  & Arrival & Leave \\
\hline
3 & John & 15/3/2012 & 30/3/2012 \\
4 & Antoon & 25/3/2012 & 4/4/2012 \\
5 & Daan & 18/3/2012 & 2/4/2012\\
3 & John & 4/4/2012 & UC \\
\hline\\
\end{tabular}
}
\end{samepage}

\begin{definition}
\label{def:current-in-relation}
\emph{Current $\left(a_k \right)$.}
Consider the elements in definition \ref{def:valid-time-relation}. The function Current $\left(a_k \right)$ returns the crisp time interval $i$ if the tuple $t$ with the primary key values given by $a_k$ is currently in the instance $r$ of the relation $R$. The function returns the emtyset otherwise.
% 
% Consider $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity and $I = \left(S_i, UC \right)$ the set of attribute defining the time interval. Let $R \subseteq \A\  x\  \I$ be a valid-time relation. The function Current $\left(R, A \right)$ returns the crisp time interval $I$ if the entity $A$ is currently in the relation $R$ or the empty set, if the entity $A$ is not currently in the relation. It is defined as follows:

%Consider an object $A$ and a relation $R$ in a temporal database. The function Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ that is current in the relation $R$, that is it, $I = \left[a, +\infty \right]$.

\begin{align}
\label{eq:current-in-relation}
\mbox{Current} \left(a_k\right) &=& \\ 
\begin{cases}
\nonumber
t & \mbox{ if } \exists t \in r : i = \left(s, UC \right)\\
\emptyset & \mbox{ otherwise}
\end{cases}
\end{align}
\end{definition}

For example, the function Current(R, (3,`John')) returns the time interval $I = \left(4/4/2012 , UC  \right)$. Conversely, Current(R, (4,`Antoon')) returns the empty set.

The Allen's relations between two time intervals are shown in Figure \ref{fig:allen-rel}. The complete implementation of the relations using only the starting and the ending points is defined in~\cite{Nagypal2003}.

\begin{samepage}
\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=0.5]{./graphs/allen.pdf}

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:allen-rel}Allen's relations for two crisp time intervals $I$ and $J$. Note that the crisp time interval $I$ is fixed and the different positions of the interval $J$ illustrate the Allen's relations.}
\vspace*{13pt}
\end{samepage}

We will use two of the Allen's relations: Overlaps and During which will be defined as follows.

\begin{definition}
\emph{Overlaps:}
 \label{def:overlaps}
Given two time intervals $i_1 = \left(s_1, e_1 \right)$ and $i_2 = \left(s_2 e_2\right)$, it is said that $i_1$ overlaps $i_2$ if:
\begin{equation}
 \label{eqn:overlaps}
i_1 \text{ Overlaps } i_2  = \left(s_1 < s_2  \right) \wedge \left(e_1 < e_2  \right) 
\end{equation}
\end{definition}

\begin{definition}
\emph{During:}
 \label{def:during}
Given two time intervals $i_1 = \left(s_1, e_1 \right)$ and $i_2 = \left(s_2 e_2\right)$, it is said that $i_1$ during $i_2$ if:
\begin{align}
 \label{eqn:during}
i_1\text{ During } i_2  =& \left(s_1 > s_2  \right) \wedge \left(e_1 > s_2 \right) \wedge \\
\nonumber & \left(e_1 < e_2 \right)
\end{align}
\end{definition}

In the rest of the paper, and without losing generality, we will consider that the time granularity are days. Also, the dates will be given in the format \emph{dd/mm/yyyy}.

\begin{definition}
\emph{CloseR$\left(i_1, i_2\right)$:}
\label{def:close-a-crisp-interval-r}
Consider two crisp intervals $i_1 = \left(s_1, e_1 \right)$ and $i_2 = \left(s_2, e_2\right)$. The CloseR$\left(i_1, i_2 \right)$ function allows to close the right-open interval $i_1$ with respect to the first value $s_2$ in $i_2$:
\begin{align}
\label{eq:close-a-crisp-interval}
\mbox{CloseR} \left( i_1, i_2\right) &=& \\ 
\begin{cases}
\nonumber
%I & \mbox{ if } b \neq +\infty \\
%I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
i_1 = \left(s_1,s_2 \right( & \mbox{ if } e_1 = UC \mbox{ and } i_2 \mbox{ During } i_1 \\
i_1 & \mbox{ otherwise}
\end{cases}
\end{align}
\end{definition}

For example, consider two time intervals, $i_1 = \left(4/4/2012, UC \right)$ and $i_2 = \left(24/4/2012, UC \right)$. The result of applying the CloseR$\left(i_1, i_2 \right)$ is $i_1 = \left(4/4/2012, 23/4/2012 \right)$.

@@

Now it is possible to close the current version of an entity by using \eqref{eq:close-a-crisp-interval} and \eqref{eq:current-in-relation}. This functionality is required to add or update new information about an existing entity in the relation.

\begin{definition}
\label{def:close-current-version}
\emph{Close-current$\left(r, t\right)$}.
Consider the elements in definition \ref{def:valid-time-relation}. The function Close-current$\left(r, t \right)$ closes any current version $t_k$ of the entity given by $t$ if it exists and add the new version $t$. For the implementation the following variables are defined:
\begin{align}
 \label{eq:variables-current-version}
t_{CUR} = \mbox{ Current } \left(t[A_K] \right)\\
\nonumber
t_{UP} = t_{CUR}; t_{UP}\left[i \right] = \mbox{Close }\left(t_{UP}\left[i \right], t \right)
\nonumber
\end{align}
Then, $t_{CUR}$ is the current version of the entity given by the tuple $t$, and $t_{UP}$ is the updated version of the tuple $t_{CUR}$. In this updated version, the time interval given by $i_{UP}$ is closed with respect to the tuple $t$.

\begin{eqnarray}
\label{eq:close-current}
\text{Close-current} \left(r, t \right) =\\
\begin{cases}
\nonumber
r - t_{CUR} \cup \lbrace t_{UP}\rbrace \cup \lbrace t \rbrace
& \mbox{ if } t_{CUR} \neq \emptyset \\
% R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right)\right \rbrace \cup \left \lbrace\left(A, J\right) \right \rbrace  \\
% \nonumber
% \mbox{ if  Current } \left(R, A \right) \neq \emptyset\\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
\nonumber r , \text{ otherwise}
\end{cases}
\end{eqnarray}
\end{definition}

For example, consider the patient with ID = 3 in Table \ref{tbl:library-sample}. The current version arrived on 4/4/2012. The patient left the hospital but, for some reason, the date  was not registered. The patient arrived again at the hospital on 24/4/2012. Then, the Close-current function allows to close the current version of the patient and insert a new version. First, the function CloseR is applied with $I = \left(4/4/2012, UC \right)$ and $J = \left(24/4/2012, UC \right)$. Hence, the value for the time interval $I$ is $ \left(4/4/2012, 23/4/2012 \right)$. Then, the modifications on the value for the time interval $I$ are stored. Finally, a new row with the current values of the patient and the time interval $J$ is stored.
The result of this operation is illustrated on Table \ref{tbl:library-sample-2}.

%It is also possible to append at the beggining by using \eqref{def:close-a-crisp-interval-l} and \eqref{def:first-in-relation}.
%\begin{definition}
%\label{def:append-first-version}
%Consider an object $A$, the relation $R$ and a time interval $J$. Then, the function append-first$\left(R,A,J \right)$ closes any right open version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
%
%\begin{eqnarray}
%%\label{eq:close-current}
%\text{Append-first} \left(R, A, J \right) =\\
%\begin{cases}
%\nonumber
%\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseL } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
%\nonumber
%\mbox{ if } I = \mbox{ First } \left(R, A \right) \wedge J < \mbox{ CloseL } \left(I, J \right)   \\
%\nonumber R , \text{ in any other case}
%\end{cases}
%\end{eqnarray}
%\end{definition}
\subsubsection{\label{subsubsec:modify}Modify}
This operation adds new information about an existing entity $A$ in the relation $R$. The modify operation does not remove any previous value of the object $A$. It closes the current version for the entity $A$ and adds a new version.

%Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[a, +\infty \right]$.

\begin{definition}
 \emph{Modify(R, (A,I))}.
Consider the elements in definition \ref{def:valid-time-relation}. The algorithm for the modify operation is defined as follows.
\end{definition}

\begin{align}
\label{eq:modify}
\MOD \left( r, t\right) =\\
%\begin{cases}
\nonumber
\mbox{ Close-current }\left(r, t\right) %\\
% \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
% R,  \mbox{ otherwise }
%\end{cases}
\end{align}
